---
title: "assignment_02"
author: "DIANA LIN"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 02

Below are 6 exercises to help you practice writing unit tests using the `testthat` package in R. 
You will be writing unit tests for the vectorized functions you created in assignment 1.
This assignment was designed to be completed in about 60 minutes and is deliberately designed to be shorter to accommodate the team project.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(testthat)
library(here)
library(glue)
mtcars_original <- mtcars
```

### Task 1: Test function inputs and outputs

#### 1.1 - Below is a function that computes the mean of every column in a dataframe. Use the appropriate `testhat` function to test that the function input is a dataframe.

```{r prob1}
# Your function from Assignment 1, Task 1.1
# Copied this from Assignment 1
output <- map_dbl(mtcars, mean)
unname(output)
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln1}
output <- function(df) {
  test_that("df is dataframe", {
    expect_s3_class(df, "data.frame")
  })
  
  return(map_dbl(df, mean))
}
```

This code chunk should fail if the test is written correctly, however `eval = FALSE` must be set in order to knit:
```{r soln1 check fail, eval=FALSE}
# To check you wrote the unit test properly, this code should throw an error
# Keep eval=FALSE or else cannot knit
output(1)
```

The following error output is printed when running the individual code chunk:
```
Error: Test failed: 'df is dataframe' * `df` is not an S3 object * `df` inherits from `numeric` not `data.frame`.
```

This code chunk should execute successfully, showing the mean per column:
```{r soln1 check success}
# To check you wrote the unit test properly, this code should run
# change eval=TRUE since above code chunks have been written
output(mtcars)
```

#### 1.2 - Below is a vectorized function that divides the values in each column of the `mtcars` dataset by the maximum in that column. Use the appropriate `testthat` function to check that the first value in the output is equal to 0.619469 (hint: to retrieve the first value `output[[1]][1]`) (hint: set `tolerance=1e-5`).

```{r prob2}
# Your function from Assignment 1, Task 1.2
# Copied this from Assignment 1
output <-
  as.data.frame(map_dfc(mtcars, function(df)
    df <- df / max(df, na.rm = TRUE)))
rownames(output) <- rownames(mtcars)
head(output)
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln2}
# tolerance=1e-6 so a last digit change will break the test
test_that("First value is 0.619469", {
  expect_equal(output[[1]][1], 0.619469, tolerance = 1e-6)
})
```

#### 1.3 - Below is a vectorized function that creates a ggplot with fuel efficiency on the x-axis and horsepower on the y-axis for three cylinder levels (4, 6, 8). Use the appropriate `testthat` function to check that the number of plots being outputted is equal to 3.

```{r prob3}
# Your function from Assignment 1, Task 1.3
# Copied this from Assignment 1
cylinders <- sort(unique(mtcars_original[['cyl']]))

plots <-
  map(cylinders, function(x)
    mtcars_original %>% filter(cyl == x) %>% ggplot() +
      theme_bw() +
      geom_point(aes(x = mpg, y = hp)) +
      labs(x = 'Fuel efficiency (mpg)',
           y = 'Horsepower (hp)') +
      ggtitle(glue::glue(
        "Horsepower and Fuel efficiency for ", x, " cylinders"
      )))

plots
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln3}
# check that the number of plots plotted is 3
test_that("Number of plots is 3.", {
  expect_equal(length(plots), 3)
})
```

### Task 2: Writing your own unit tests

Some ideas for writing unit tests include checking if an input or output is the correct length, type, value, etc.

#### 2.1  - Below is sample code that computes the number of unique values in each column of `mtcars` in a vectorized way.  Use the appropriate `testthat` function to check the output of this code.

```{r prob4}
output <- map(mtcars, unique) %>%
  map_dbl(length)
output
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln4}
# check that the output is the same as the number of columns in the data frame
test_that("Output length is same as number of columns", {
  expect_equal(length(output), ncol(mtcars))
})
```

#### 2.2 - Below is sample code that takes input from various columns in the `diamonds` dataset and outputs a string containing information from the input. Use the appropriate `testthat` function to write 2 tests: 1 test to check the input of the `output` function and 1 test to check the output of the `output` function .

```{r, prob5, echo = TRUE}
dmonds <- diamonds %>%
  slice(1:4)

list_of_things <-
  list(seq(nrow(dmonds)), dmonds$price, dmonds$carat)

output <- function(input_list) {
  pmap_chr(input_list,
           ~ paste0("Diamond #", ..1 ,
                    " sold for $", ..2,
                    " and was ", ..3, " carats"))
}
output(list_of_things)
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln5}
# since pmap takes a list, check if list
test_that("Input list is a list", {
  expect_type(list_of_things, "list")
})

# since output should be string, check character vector
test_that("Output is character vector.", {
  expect_type(output(list_of_things), "character")
})
```

#### 2.3 - Below is sample code that splits the iris dataset into three species, and saves each dataframe as a separate file. Use the appropriate `testthat` function to check that these files were successfully created. Bonus points if you use the appropriate `purrr` function to check if all the files were successfully created. (hint: `file.exists(here::here(glue::glue("1", ".csv")))` returns `TRUE` if the file exists and `FALSE` if it doesn't)

```{r prob6}
iris %>%
  nest(data = -Species) %>%
  pmap( ~ write_csv(.y, path = here::here(glue::glue(.x, ".csv"))))
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln6 }
# run same initial code, but instead of writing csv, check that those csv's exist
iris %>%
  nest(data = -Species) %>%
  pmap( ~ test_that("Written csv files exist.", {
    expect_true(file.exists(here::here(glue::glue(., ".csv"))))
  }))
```

## Additional Resources

1. [Functions from the `testthat` package:](https://testthat.r-lib.org/reference/index.html)
2. [Testing chapter from the `R-packages` book:](https://r-pkgs.org/tests.html)
