---
title: "assignment_01"
author: "DIANA LIN"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 01

Below are 6 exercises to help you practice vectorizing for-loops, and converting for-loops into vectorized expressions using `purrr` functions.
This assignment was designed to be completed in about 60 minutes and is deliberately designed to be shorter to accommodate the team project.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
mtcars_original <- mtcars
```
### Task 1: Convert for-loop code to vectorized expressions using `purrr` functions

#### 1.1 - Below is sample code that computes the mean of every column in the `mtcars` dataset. Use the appropriate `purrr` function to vectorize this task.

```{r prob1}
output <- vector("double", ncol(mtcars))
for (c in seq_along(mtcars)){
  output[[c]] <- mean(mtcars[[c]])
}
output
```

Your solution to vectorize the code above (using a `purrr` function):

```{r soln1}
# use map_dbl to return a double vector
output <- map_dbl(mtcars, mean)
unname(output) # unname to remove names in vector to match output above
```

#### 1.2 - Below is sample code that divides the values in each column of the `mtcars` dataset by the maximum in that column. Use the appropriate `purrr` function to vectorize this task.

```{r prob2}
for (c in seq_along(mtcars)){
  mtcars[[c]] <- mtcars[[c]] / max(mtcars[[c]], na.rm = TRUE)
}
head(mtcars)
```

Your solution to vectorize the code above (using a `purrr` function):

```{r soln2}
output <- as.data.frame(map_dfc(mtcars, function(df) df <- df / max(df, na.rm = TRUE))) # convert to data frame because row names are now allowed on a tibble
rownames(output) <- rownames(mtcars) # add rownames so it looks like it does in the above chunk output
head(output) # head to reduce output as above
```

#### 1.3 - Below is sample code that creates a ggplot with fuel efficiency on the x-axis and horsepower on the y-axis for three cylinder levels (4, 6, 8). Use the appropriate `purrr` function to vectorize this task.

```{r prob3}
cylinders <- sort(unique(mtcars_original[['cyl']]))
plots <- vector("list",length(cylinders))

for (d in 1:length(cylinders)){
  plots[[d]] <- mtcars_original %>% 
    filter(cyl == cylinders[d]) %>%
    ggplot() + 
    theme_bw() +
    geom_point(aes(x = mpg, y = hp) ) + 
    labs(x = 'Fuel efficiency (mpg)',
         y = 'Horsepower (hp)') + 
    ggtitle(glue::glue("Horsepower and Fuel efficiency for ", cylinders[d], " cylinders"))
}
plots
```

Your solution to vectorize the code above (using a `purrr` function):

```{r soln3}
# get cylinder vector (as above)
cylinders <- sort(unique(mtcars_original[['cyl']]))

# use function(x) to customize dplyr/ggplot combination of functions
plots <-
  map(cylinders, function(x)
    mtcars_original %>% filter(cyl == x) %>% ggplot() +
      theme_bw() +
      geom_point(aes(x = mpg, y = hp)) +
      labs(x = 'Fuel efficiency (mpg)',
           y = 'Horsepower (hp)') +
      ggtitle(glue::glue(
        "Horsepower and Fuel efficiency for ", x, " cylinders"
      )))

plots
```

### Task 2: Convert vectorized expressions into one that uses for-loops

#### 2.1  - Below is sample code that computes the number of unique values in each column of `mtcars` in a vectorized way. Turn this vectorized expression into one that uses for-loop(s) that reproduces the output (numbers should be the same, the format can be different).
```{r prob4}
map(mtcars_original, unique) %>% 
  map_dbl(length)
```

Your solution to "un-vectorize" the code above using for-loops:

```{r soln4}
# allot memory for output
vectorized <- vector("double",length(mtcars_original))

# iterate through the columns, unique and then get length
for (i in seq_along(mtcars_original)) {
  vectorized[[i]] <- unique(mtcars_original[[i]]) %>%
    length()
}
names(vectorized) <- names(mtcars_original)
vectorized
```

#### 2.2 - Below is sample code that takes input from various columns in the `diamonds` dataset and outputs a string containing information from the input. Turn this vectorized expression into one that uses for-loop(s) that reproduces the output (numbers should be the same, the format can be different).

*Hint:* Note the use of `pmap_chr` to iterate over three columns (diamond number, price and carat).

```{r, prob5, echo = TRUE}
dmonds <- diamonds %>% 
  slice(1:4)

list_of_things <- list(seq(nrow(dmonds)), dmonds$price, dmonds$carat)
output <- pmap_chr( list_of_things,
                   ~paste0("Diamond #", ..1 ,
                           " sold for $",..2,
                           " and was ",..3, " carats"))
output
```

Your solution to "un-vectorize" the code above using for-loops:

```{r soln5}
library(glue)
# add solution here
dmonds <- diamonds %>% 
  slice(1:4)

# allot the memory for output
output <- vector("character", nrow(dmonds))

# get the vectors to iterate through
num_diamonds <- seq(nrow(dmonds))
price_diamonds <- dmonds$price
carat_diamonds <- dmonds$carat

# iterate through all three using the same index i for each vector
for (i in seq_along(seq(nrow(dmonds)))) {
  output[[i]] <- glue("Diamond #{num_diamonds[i]} sold for ${price_diamonds[i]} and was {carat_diamonds[i]} carats")
}

output
```

#### 2.3 - Below is sample code that splits the iris dataset into three species, and saves each dataframe as a separate file. Turn this vectorized expression into one that uses a for-loop that reproduces the output.

*Hint 1*: `map2(x, y, f)` is the same as `pmap(list(x, y), f)`

*Hint 2*: Multiple files can be saved in a vectorized way using the`pmap` function. The [`nest` function](https://tidyr.tidyverse.org/reference/nest.html) is used to create a list-column of data frames for each species from the `iris` dataset.

```{r prob6}
iris %>%
  nest(data = -Species) %>%
  pmap(~write_csv(.y, path = here::here(glue::glue(.x, ".csv"))))
```

Your solution to "un-vectorize" the code above using for-loops:

```{r soln6}
# get the three species
species <- sort(unique(iris$Species))

# allot the memory
dfs <- vector("list", length(species))

# iterate through each species and write the csv and put into dfs
for (i in seq_along(species)) {
  dfs[[i]] <- iris %>%
    filter(Species == species[i]) %>% # filter for specific species
    select(-Species) %>% # remove species from the data frame
    as_tibble() # convert to tibble as above output was a tibble
  
  write_csv(dfs[[i]], path = here::here(paste0(i, ".csv"))) # write to current working directory using the same naming convention as above
}

dfs # print tibble to see
```

## Additional Resources

1. Chapter 21 of [R for Data Science](https://r4ds.had.co.nz/iteration.html).
2. [Learn to purr](http://www.rebeccabarter.com/blog/2019-08-19_purrr/) blog post.
3. Chapter 9 of [Advanced R for Data Science](https://adv-r.hadley.nz/functionals.html).